---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---


```{r}
library(data.table)
library(caret)
library(MLeval)
library(glmnet)
library(ROCR)
```

```{r}
mci <- fread('mci.csv')
proteins <- fread('proteins_for_ML.csv')
Olink_submitted_full <- fread('Olink_submitted_full_for_ML.csv')
tT1.sfE <- fread('limma-MCI-no-duplicates-annotated.csv')

# Abeta and ptau data 

abeta <- fread('Ab40-42 results (10.25.2021).csv')
abeta[`AB1-40` == 'Above Highest Calibrator', `AB1-40` := 278]
abeta[`AB1-42` == 'Below Lowest Calibrator', `AB1-42` := 8.8]
abeta <- abeta[1:101]
abeta[,`AB1-40` := as.numeric(`AB1-40`)]
abeta[,`AB1-42` := as.numeric(`AB1-42`)]
abeta[,ABeta_Ratio := `AB1-42`/`AB1-40`]

ptau <- fread('olink_ptau181_report_092821.csv')

biomarkers_prognosis <- tT1.sfE[tT1.sfE$P.Value < 0.05,]$Assay.1
cols<-c("Assay", "PIDN", "category_steve",biomarkers_prognosis)
dat<-proteins[Assay %in% mci$Plasma_Sample, .SD,.SDcols=cols]

dat[Olink_submitted_full, on=.(Assay = Plasma_Sample), `:=` ( Age_at_Visit = i.Age_at_Visit,  Global_CDR =i.Global_CDR, CDR_Sum_of_Boxes = i.CDR_Sum_of_Boxes, FAQ_Total = i.FAQ_Total, MMSE = i.MMSE, Sex = i.Sex, Education =i.Education )]

dat[, Sex := ifelse(Sex=="Male", 0, 1)]
dat[, Education :=ifelse(Education=="12 Years",1,
                  ifelse(Education=="13-15 Years",2,
                  ifelse(Education=="16 Years",3,4)))]

dat[Olink_submitted_full, on=.(PIDN), APOE4 := i.APOE4]
dat[, APOE4 := as.numeric(APOE4)]
dat[,Phenotype := factor(category_steve, levels=c("MCI_Stable_AD","MCI_Decline_AD"))]

dat[abeta,on=.(Assay = SampleID),ABeta_Ratio := i.ABeta_Ratio]
#mean value of ratio 
dat[Assay == 'H0401.1',ABeta_Ratio := 0.1245184]

dat[ptau,on=.(Assay = SampleID), ptau := `i.pTau181 (pg/mL)`]
dat[Assay == 'H0401.1',ptau := 2.827273]

# Join with MSD pTau
# msd<-fread("mci_with_ptau.csv")
# dat[msd, on=.(Assay = Plasma_Sample), ptau_high := i.ptau_high]

#fwrite(dat,"mci-all-dat.csv")
# remove vars that don't go into ML
dat[, c("PIDN","Assay","category_steve") := NULL]


# !!!!
normalize = preProcess(dat, method = c("center", "scale", "zv"), verbose = TRUE)
dat = predict(normalize, dat)
dat[is.na(dat)] <-0.001
n=dim(dat)[2]
```

SVM via caret 

```{r}
# set.seed(9)
# tC = trainControl(method = "cv", number = 10, savePredictions = TRUE, classProbs = TRUE, verboseIter = TRUE)
# 
# modelSVM = train(Phenotype ~ ., dat, method = "svmLinear2", trControl = tC)
# confusionMatrix(modelSVM)
# res<-evalm(modelSVM)
# res
# save(modelSVM, file=paste0("modelSVM_",Sys.Date(),".Rdata"))
# plot(varImp(modelSVM), main = "Variable Importance", cex.lab = 0.1, n.var=10)


```



Lasso via glment 

```{r}
# set.seed(5281995)
# f <- as.formula(Phenotype ~ .)
# 
# # split, 80 percent train/CV and 20 percent test
# train.index <- createDataPartition(dat$Phenotype, p = .8, list = FALSE)
# train_data <- dat[ train.index,]
# test_data  <- dat[-train.index,]
# 
# train_x <- model.matrix(f, train_data)
# train_y <- as.factor(ifelse(as.matrix(train_data[,52]) == 'MCI_Decline_AD',1,0))
# 
# # not enough data for auc, uses deviance
# mci_glmnet <- cv.glmnet(x = train_x, y = train_y,family = "binomial",type.measure = "auc", nfolds = 5)
# 
# plot(mci_glmnet)
# exp(coef(mci_glmnet,s='lambda.min'))
# 
# saveRDS(mci_glmnet, file= 'glmnet_44_proteins.rds')
# 
# 
# test_x <- model.matrix(f, test_data)
# test_y <- as.factor(ifelse(as.matrix(test_data[,52]) == 'MCI_Decline_AD',1,0))
# 
# results <- data.table(Prediction = predict(mci_glmnet, newx = test_x, s = "lambda.min", type = "class")[,1],Target = test_y)
# 
# results[,Correct := ifelse(Prediction == Target,1,0)]
# results[,Predicted_Prob := predict(mci_glmnet, newx = test_x, s = "lambda.min", type = "response")[,1]]
# results[,sum(Correct)]
# 
# 
# AUC <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'auc')
# AUC@y.values
# 
# perf <- performance(prediction(results$Predicted_Prob,results$Target), measure = "tpr", x.measure = "fpr")
# plot(perf,colorize = T)
# lines(x = c(0,100), y = c(0,100))

```
Glmnet, but put all data into model matrix 

```{r}
set.seed(5281995)
f <- as.formula(Phenotype ~ .)

x <- model.matrix(f, dat)
y <- as.factor(ifelse(as.matrix(dat[,52]) == 'MCI_Decline_AD',1,0))
# x = x[, -1] # why do we have two intercepts?

# not enough data for auc, uses deviance
mci_glmnet <- cv.glmnet(x = x, y = y,family = "binomial",type.measure = "auc", nfolds = 5, keep=T)


#saveRDS(mci_glmnet,file = 'full_model_abeta_plasma_glmnet_v2.rds')
#saveRDS(x,file = 'full_model_abeta_plasma_x_matrix_v2.rds')


plot(mci_glmnet)
exp(coef(mci_glmnet,s='lambda.min'))

#saveRDS(mci_glmnet, file= 'glmnet_44_proteins.rds')


idmin = match(mci_glmnet$lambda.min, mci_glmnet$lambda)
plot(roc.glmnet(mci_glmnet$fit.preval, newy = y)[[idmin]])
roc_output <- roc.glmnet(mci_glmnet$fit.preval, newy = y)[[idmin]]
setDT(roc_output)
roc_output[,TP := TPR * 30]
roc_output[,TN := (1 - FPR) * 30]
roc_output[,ACC := (TP + TN) / 60]
roc_output[,PPV := TP/30]
roc_output[,NPV := TN /30]

# this is full model output 
fwrite(roc_output,file='ROC_coordinates_full_model_with_ptau_abeta_idmin_29.csv')


# ugly, SD/AN to update 
# ROC_plot <- ggplot(roc_output, aes(x=FPR, y=TPR)) +
#   geom_point() + 
#   geom_abline(intercept = 0, slope = 1) + 
#   ggtitle('ROC Plot MCI GLMnet Including Biomarkers') + 
#   geom_text(aes(label=ifelse(FPR == 0 & TPR == 0.7,'FPR == .7','')),hjust = , vjust= )
# ROC_plot

# results <- data.table(Prediction = predict(mci_glmnet, newx = x, s = "lambda.min", type = "class")[,1],Target = y)
# results[,Correct := ifelse(Prediction == Target,1,0)]
# results[,Predicted_Prob := predict(mci_glmnet, newx = x, s = "lambda.min", type = "response")[,1]]
# results[,sum(Correct)]
# 
# cm = confusionMatrix(results$Prediction,results$Target)


#AUC <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'auc')
#AUC@y.values

# auc<-ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'tpr','fpr')
# plot(auc)

# perf <- performance(prediction(results$Predicted_Prob,results$Target), measure = "tpr", x.measure = "fpr")
# plot(perf,colorize = T)
# lines(x = c(0,100), y = c(0,100))

# lamdba investigation 


lambda_mci_glmnet <- data.table(lambda = mci_glmnet$lambda, 
                                cv_mean = mci_glmnet$cvm,
                                cv_sd = mci_glmnet$cvsd,
                                cv_lo = mci_glmnet$cvlo,
                                cv_up = mci_glmnet$cvup)

View(lambda_mci_glmnet[lambda == mci_glmnet$lambda.min])

fwrite(lambda_mci_glmnet,file='lambda_sequence_mci_glmnet_full_model_with_ptau_abeta_idmin_29.csv')
mci_glmnet_coef <- data.table(var = names(coef(mci_glmnet,s='lambda.min')[,1]),coef = exp(coef(mci_glmnet,s='lambda.min')[,1]))

#fwrite(mci_glmnet_coef,file='coef_mci_glmnet_full_model_with_abeta_ptau_v2.csv')
```


Baseline model plus abeta/ptau 

```{r}
# need to remove plasma biomarkers 
set.seed(5281995)
f <- as.formula(Phenotype ~ .)

cols <- c("Age_at_Visit","CDR_Sum_of_Boxes","FAQ_Total","MMSE","Sex","Education","APOE4","Phenotype","ABeta_Ratio","ptau")

dat_baseline_abeta_ptau <- dat[,..cols]

x <- model.matrix(f, dat_baseline_abeta_ptau)
y <- as.factor(ifelse(as.matrix(dat_baseline_abeta_ptau[,8]) == 'MCI_Decline_AD',1,0))

mci_glmnet <- cv.glmnet(x = x, y = y,family = "binomial",type.measure = "auc", nfolds = 5, keep=T)

plot(mci_glmnet)
exp(coef(mci_glmnet,s='lambda.min'))

saveRDS(mci_glmnet, file= 'glmnet_baseline_plus_abeta_ptau.rds')


idmin = match(mci_glmnet$lambda.min, mci_glmnet$lambda)
plot(roc.glmnet(mci_glmnet$fit.preval, newy = y)[[idmin]])
roc_output <- roc.glmnet(mci_glmnet$fit.preval, newy = y)[[idmin]]
setDT(roc_output)
roc_output[,TP := TPR * 30]
roc_output[,TN := (1 - FPR) * 30]
roc_output[,ACC := (TP + TN) / 60]
roc_output[,PPV := TP/30]
roc_output[,NPV := TN /30]

# this is full model output 
fwrite(roc_output,file='ROC_coordinates_baseline_with_ptau_abeta_idmin_CGM.csv')

lambda_mci_glmnet <- data.table(lambda = mci_glmnet$lambda, 
                                cv_mean = mci_glmnet$cvm,
                                cv_sd = mci_glmnet$cvsd,
                                cv_lo = mci_glmnet$cvlo,
                                cv_up = mci_glmnet$cvup)

View(lambda_mci_glmnet[lambda == mci_glmnet$lambda.min])

fwrite(lambda_mci_glmnet,file='lambda_sequence_mci_glmnet_baseline_with_ptau_abeta_idmin_CGM.csv')

```


rerun glmnet on baseline model (no proteins)

```{r}
set.seed(12345)
f <- as.formula(Phenotype ~ .)

baseline_dat <- dat[,45:ncol(dat)]

x <- model.matrix(f, baseline_dat)
y <- as.factor(ifelse(as.matrix(baseline_dat[,8]) == 'MCI_Decline_AD',1,0))


mci_baseline_glmnet <- cv.glmnet(x = x, y = y,family = "binomial",type.measure = "auc", nfolds = 5,keep = T)

plot(mci_baseline_glmnet)
exp(coef(mci_baseline_glmnet,s='lambda.min'))

idmin_baseline = match(mci_baseline_glmnet$lambda.min, mci_baseline_glmnet$lambda)
roc_baseline_output <- roc.glmnet(mci_baseline_glmnet$fit.preval, newy = y)[[idmin_baseline]]
setDT(roc_baseline_output)
roc_baseline_output[,TP := TPR * 30]
roc_baseline_output[,TN := (1 - FPR) * 30]
roc_baseline_output[,ACC := (TP + TN) / 60]
roc_baseline_output[,PPV := TP/30]
roc_baseline_output[,NPV := TN /30]

fwrite(roc_baseline_output,file='ROC_coordinates_baseline_model.csv')

#saveRDS(mci_glmnet, file= 'glmnet_44_proteins.rds')

# results <- data.table(Prediction = predict(mci_baseline_glmnet, newx = x, s = "lambda.min", type = "class")[,1],Target = y)
# 
# results[,Correct := ifelse(Prediction == Target,1,0)]
# results[,Predicted_Prob := predict(mci_baseline_glmnet, newx = x, s = "lambda.min", type = "response")[,1]]
# results[,sum(Correct)]
# 
# 
# AUC <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'auc')
# AUC@y.values
# 
# perf <- performance(prediction(results$Predicted_Prob,results$Target), measure = "tpr", x.measure = "fpr")
# plot(perf,colorize = T)
# lines(x = c(0,100), y = c(0,100))

# lamdba investigation 


lambda_mci_baseline_glmnet <- data.table(lambda = mci_baseline_glmnet$lambda, 
                                cv_mean = mci_baseline_glmnet$cvm,
                                cv_sd = mci_baseline_glmnet$cvsd,
                                cv_lo = mci_baseline_glmnet$cvlo,
                                cv_up = mci_baseline_glmnet$cvup)

View(lambda_mci_baseline_glmnet[lambda == mci_baseline_glmnet$lambda.min])


fwrite(lambda_mci_baseline_glmnet,file='lambda_sequence_mci_baseline_glmnet.csv')
mci_baseline_glmnet_coef <- data.table(var = names(coef(mci_baseline_glmnet,s='lambda.min')[,1]),coef = exp(coef(mci_baseline_glmnet,s='lambda.min')[,1]))

fwrite(mci_baseline_glmnet_coef,file='coef_mci_baseline_glmnet.csv')
```



# eval on the test/replication set 



```{r}
# make test protein data matrix, need to add in ptau/abeta

proteins_test<-read.csv("Olink_SVM_Test.csv", header=T,stringsAsFactors = F)
setDT(proteins_test)
proteins_test[,2:369] <- map_dfc(proteins_test[,2:369],~as.double(.x))

final_samples <- fread('2020-06-25-Blood-Visits-Scans.csv')

proteins_test <- merge(proteins_test,final_samples[,.(DrawID,FAQ_Total,MMSE,CDR_Sum_of_Boxes)],by.x = "Assay",by.y = "DrawID",all.x = T)

outcomes <- fread('Final sample selection_MCI validation cohort.csv')

proteins_test[outcomes,on=.(Assay = Assay.1),Phenotype := i.Cog_Status]
#View(proteins_test[,..biomarkers_prognosis])


outcomes[,'Age_at_Visit' := `Age at blood draw`]
outcomes[,Sex := V7]
outcomes[, Sex := ifelse(Sex=="M", 0, 1)]
outcomes[,MMSE := as.numeric(MMSE)]
outcomes[APOE %in% c("e3/e4", "e4/e2", "e4/e4"), APOE4 := 1]
outcomes[APOE %in% c("e3/e3", "e3/e2","e2/e2"), APOE4 := 0]
outcomes[, Education :=ifelse(Education=="12 Years",1,
                  ifelse(Education=="13-15 Years",2,
                  ifelse(Education=="16 Years",3,4)))]

proteins_test[outcomes,on=.(Assay = Assay.1),`:=` 
              (Age_at_Visit = i.Age_at_Visit,
                CDR_Sum_of_Boxes = i.CDR_Sum_of_Boxes,
                MMSE = i.MMSE,
                Sex = i.Sex,
                Education = i.Education,
                APOE4 = i.APOE4)]
#fwrite(proteins_test,"proteins_validation_set.csv")
# filter for only proteins of interest 
test_model_matrix <- proteins_test[,.SD,.SDcols =   c(biomarkers_prognosis,'Age_at_Visit',
'CDR_Sum_of_Boxes','FAQ_Total',                                               'MMSE','Sex','Education','APOE4','Phenotype')]


setcolorder(test_model_matrix,colnames(dat))

# QC this, need to remove NAs
test_model_matrix <- test_model_matrix[!is.na(Age_at_Visit)]

# normalize, apply training preprocessing??????? 
test_model_matrix <- predict(normalize, test_model_matrix)


set.seed(5281995)
f <- as.formula(Phenotype ~ .)

# model 1 
x <- model.matrix(f, test_model_matrix)
y <- as.factor(ifelse(as.matrix(test_model_matrix[,52]) == 'MCI Decline',1,0))

#mci_glmnet <- load()

results <- data.table(Prediction = predict(mci_glmnet, newx = x, s = "lambda.min", type = "class")[,1],Target = y)

results[,Correct := ifelse(Prediction == Target,1,0)]
results[,Predicted_Prob := predict(mci_glmnet, newx = x, s = "lambda.min", type = "response")[,1]]
results[,sum(Correct)]

# results from model 1 
AUC <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'auc')
AUC@y.values

# [[1]]
# [1] 0.675737


perf <- performance(prediction(results$Predicted_Prob,results$Target), measure = "tpr", x.measure = "fpr")
plot(perf,colorize = T)
lines(x = c(0,100), y = c(0,100))

fwrite(results,file='test_replication_set_results.csv')


acc <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'acc')
acc@y.values[[1]][max_acc_idx]

max_acc_idx <- which.max(acc@y.values[[1]]) # 28 

sens <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'sens')
sens@y.values[[1]][max_acc_idx]  # 0.8571429

spec <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'spec')
spec@y.values[[1]][max_acc_idx] # 0.5714286

fpr <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'fpr')
fpr@y.values[[1]][max_acc_idx]  # 0.4285714

tpr <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'tpr')
tpr@y.values[[1]][max_acc_idx] # 0.8571429

tnr <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'tnr')
tnr@y.values[[1]][max_acc_idx] # 0.8571429

replication_set_metrics <- data.table(AUC = AUC@y.values,
                                      ACC = acc@y.values[[1]][max_acc_idx],
                                      SENS = sens@y.values[[1]][max_acc_idx],
                                      SPEC = spec@y.values[[1]][max_acc_idx],
                                      FPR = fpr@y.values[[1]][max_acc_idx],
                                      TPR = tpr@y.values[[1]][max_acc_idx],
                                      TNR = tnr@y.values[[1]][max_acc_idx]
                                     )

replication_set_metrics[, FP := FPR * 21]
replication_set_metrics[, TP := TPR * 21]
replication_set_metrics[, TN := TNR * 21]
replication_set_metrics[, FN := 21 - TP]

```

Replicate Table 1 on validation (replication) cohort

```{r}
# N, mean age, sd age, percent female by category 

table_1 <- proteins_test[3:nrow(proteins_test),][,.(N = .N, Age_Mean = mean(Age_at_Visit), Age_SD = sd(Age_at_Visit),Percent_Female = sum(Sex)/.N),by=.(Phenotype)]
```



Eval on test set with full model (plasma biomarkers + ptau and abeta + baseline). CAUTION: ENSURE THAT YOU EITHER EXPLICITLY LOAD IN CORRECT GLMNET MODEL (ONE WITH ALL DATA) BEFORE RUNNING THIS SECTION OR THAT MOST RECENT mci_glmnet THAT YOU CREATED WAS THE 'FULL' VERSION WITH ALL VARIABLES. 

```{r}
# make test protein data matrix, need to add in ptau/abeta

proteins_test<-read.csv("Olink_SVM_Test.csv", header=T,stringsAsFactors = F)
setDT(proteins_test)
proteins_test[,2:369] <- map_dfc(proteins_test[,2:369],~as.double(.x))

final_samples <- fread('2020-06-25-Blood-Visits-Scans.csv')

proteins_test <- merge(proteins_test,final_samples[,.(DrawID,FAQ_Total,MMSE,CDR_Sum_of_Boxes)],by.x = "Assay",by.y = "DrawID",all.x = T)

outcomes <- fread('Final sample selection_MCI validation cohort.csv')

proteins_test[outcomes,on=.(Assay = Assay.1),Phenotype := i.Cog_Status]
#View(proteins_test[,..biomarkers_prognosis])


outcomes[,'Age_at_Visit' := `Age at blood draw`]
outcomes[,Sex := V7]
outcomes[, Sex := ifelse(Sex=="M", 0, 1)]
outcomes[,MMSE := as.numeric(MMSE)]
outcomes[APOE %in% c("e3/e4", "e4/e2", "e4/e4"), APOE4 := 1]
outcomes[APOE %in% c("e3/e3", "e3/e2","e2/e2"), APOE4 := 0]
outcomes[, Education :=ifelse(Education=="12 Years",1,
                  ifelse(Education=="13-15 Years",2,
                  ifelse(Education=="16 Years",3,4)))]

proteins_test[outcomes,on=.(Assay = Assay.1),`:=` 
              (Age_at_Visit = i.Age_at_Visit,
                CDR_Sum_of_Boxes = i.CDR_Sum_of_Boxes,
                MMSE = i.MMSE,
                Sex = i.Sex,
                Education = i.Education,
                APOE4 = i.APOE4)]

proteins_test[abeta,on=.(Assay = SampleID),ABeta_Ratio := i.ABeta_Ratio]
proteins_test[ptau,on=.(Assay = SampleID),ptau := `i.pTau181 (pg/mL)`]


#fwrite(proteins_test,"proteins_validation_set.csv")
# filter for only proteins of interest 
test_model_matrix <- proteins_test[,.SD,.SDcols =   c(biomarkers_prognosis,'Age_at_Visit',
'CDR_Sum_of_Boxes','FAQ_Total',                                               'MMSE','Sex','Education','APOE4','Phenotype','ABeta_Ratio','ptau')]

setcolorder(test_model_matrix,colnames(dat))

# QC this, need to remove NAs
test_model_matrix <- test_model_matrix[!is.na(Age_at_Visit)]

#fwrite(test_model_matrix,file = 'test_set_all_data.csv')

# normalize, apply training preprocessing??????? 
test_model_matrix <- predict(normalize, test_model_matrix)


set.seed(5281995)
f <- as.formula(Phenotype ~ .)

# model 1 
x <- model.matrix(f, test_model_matrix)
y <- as.factor(ifelse(as.matrix(test_model_matrix[,52]) == 'MCI Decline',1,0))

#mci_glmnet <- load()

results <- data.table(Prediction = predict(mci_glmnet, newx = x, s = "lambda.min", type = "class")[,1],Target = y)


results[,Correct := ifelse(Prediction == Target,1,0)]
results[,Predicted_Prob := predict(mci_glmnet, newx = x, s = "lambda.min", type = "response")[,1]]
results[,sum(Correct)]

# results from model 1 
AUC <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'auc')
AUC@y.values

# [[1]]
# [1] 0.689


perf <- performance(prediction(results$Predicted_Prob,results$Target), measure = "tpr", x.measure = "fpr")
plot(perf,colorize = T)
lines(x = c(0,100), y = c(0,100))

fwrite(results,file='test_replication_set_results.csv')


acc <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'acc')
acc@y.values[[1]][which.max(acc@y.values[[1]])]

max_acc_idx <- which.max(acc@y.values[[1]]) # 28 

sens <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'sens')
sens@y.values[[1]][max_acc_idx]  # 0.8571429

spec <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'spec')
spec@y.values[[1]][max_acc_idx] # 0.5714286

fpr <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'fpr')
fpr@y.values[[1]][max_acc_idx]  # 0.4285714

tpr <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'tpr')
tpr@y.values[[1]][max_acc_idx] # 0.8571429

tnr <- ROCR::performance(prediction(results$Predicted_Prob,results$Target),measure = 'tnr')
tnr@y.values[[1]][max_acc_idx] # 0.8571429

replication_set_metrics <- data.table(AUC = AUC@y.values,
                                      ACC = acc@y.values[[1]][max_acc_idx],
                                      SENS = sens@y.values[[1]][max_acc_idx],
                                      SPEC = spec@y.values[[1]][max_acc_idx],
                                      FPR = fpr@y.values[[1]][max_acc_idx],
                                      TPR = tpr@y.values[[1]][max_acc_idx],
                                      TNR = tnr@y.values[[1]][max_acc_idx]
                                     )

fwrite(replication_set_metrics,file='replication_set_metrics_full_model_CGM.csv')

replication_set_metrics[, FP := FPR * 21]
replication_set_metrics[, TP := TPR * 21]
replication_set_metrics[, TN := TNR * 21]
replication_set_metrics[, FN := 21 - TP]

```


get beta*val for all patients model 3 (full)

```{r}
mci_glmnet <- readRDS('full_model_abeta_plasma_glmnet.rds')
x <- readRDS('full_model_abeta_plasma_x_matrix.rds')

exp(coef(mci_glmnet,s='lambda.min'))

# need to drop second beta since it's a redundant intercept 
scores <- x %*% diag(as.numeric(coef(mci_glmnet,s='lambda.min'))[-2])
scores <- as.data.table(as.data.frame(scores))

# fix names, drop column 52=Phenotype
new_names <- c('Intercept',colnames(dat)[c(1:51,53:54)])
setnames(scores,new_names)

# add phenotype back
# can safely assume dat and vdata are in the same order:
# sum(dat$Phenotype == vdata$Phenotype)
vdata <-fread("mci-all-dat.csv")
vdata[, pTau_Q := ifelse(ptau > 2.6, "T+", "T-")]
vdata[ , pTau_Q :=as.factor(pTau_Q)]
vdata[, .N, by=.(Phenotype, pTau_Q)]

scores <- cbind(scores, vdata[, .(Phenotype, raw_ptau = ptau, pTau_Q)])
# setnames(scores,'V2','Phenotype')
#names(scores)[57] = "raw_ptau"

#fwrite(scores,file = 'score_input_data_v2.csv')

```

parallel coordinates plot
https://www.data-to-viz.com/caveat/spider.html

```{r}

# load library
library(GGally)
library(viridis)
library(patchwork)
library(hrbrthemes)
library(fmsb)
# library(colormap)

# read data again
scores = fread("score_input_data_v2.csv")
pc_cols = c("IL8.1", "TIMP4", "MMP7", "TGF.alpha", "ST2", "NEFL")
which(colnames(scores) %in% pc_cols)

# check wrong scores
# sum(scores[1, .SD, .SDcols = names(scores)[2:45]] == 0)
# length(2:45)
# 44-37

# coerce to factor
scores[, pTau_Q := factor(pTau_Q, levels = c("T-", "T+"))]

scores[, Phenotype := factor(Phenotype, levels = c("MCI_Stable_AD", "MCI_Decline_AD"), labels = c("Stable", "Progressors"))]

# merge with ptau data

# create parallel coordinate plot
scores %>% .[raw_ptau <= 2.6] %>% as.data.frame() %>%
  ggparcoord(
    columns = pc_cols, groupColumn = "Phenotype", order = "anyClass",
    showPoints = TRUE,
    alphaLines = 0.3
    ) + 
  labs(y = "Value") +
  theme_bw() +
  theme(axis.text.x = element_text(face="bold", color = "black", size = 12),
        axis.title.y = element_text(face="bold", size = 12),
        axis.title.x = element_blank(),
        legend.title = element_text(face = "bold", size = 12),
        legend.position = "bottom")

## High pTau
scores %>% .[raw_ptau > 2.6] %>% as.data.frame() %>%
  ggparcoord(
    columns = pc_cols, groupColumn = "Phenotype", order = "anyClass",
    showPoints = TRUE,
    alphaLines = 0.3
    ) + 
  labs(y = "Value") +
  theme_bw() +
  theme(axis.text.x = element_text(face="bold", color = "black", size = 12),
        axis.title.y = element_text(face="bold", size = 12),
        axis.title.x = element_blank(),
        legend.title = element_text(face = "bold", size = 12),
        legend.position = "bottom")


# progressors

scores[Phenotype == "Progressors"] %>% as.data.frame() %>%
  ggparcoord(
    columns = pc_cols, groupColumn = "pTau_Q", order = "anyClass",
    showPoints = TRUE,
    alphaLines = 0.3
    ) + 
  labs(y = "Value") +
  theme_bw() +
  theme(axis.text.x = element_text(face="bold", color = "black", size = 12),
        axis.title.y = element_text(face="bold", size = 12),
        axis.title.x = element_blank(),
        legend.title = element_text(face = "bold", size = 12),
        legend.position = "bottom")



# BTH's hypothesis: for the low ptau people, their decline can be explained by the Olink biomarkers? so perhaps if tau is down then IL8 (for example) is up? etc.
# SO, for the people who are negative, do progressors have higher Olink markers compared to stable?
scores %>% as.data.frame() %>%
  ggparcoord(
    columns = pc_cols, groupColumn = "Phenotype", order = "anyClass",
    showPoints = TRUE,
    alphaLines = 0.3
    ) + 
  labs(y = "Value") +
  theme_bw() +
  theme(axis.text.x = element_text(face="bold", color = "black", size = 12),
        axis.title.y = element_text(face="bold", size = 12),
        axis.title.x = element_blank(),
        legend.title = element_text(face = "bold", size = 12),
        legend.position = "bottom")


```

ggplots BTH ptau analysis

```{r}
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(ggpubr)

vdata <-fread("mci-all-dat.csv")

vdata[, pTau_Q := ifelse(ptau > 2.6, "T+", "T-")]
vdata[ , pTau_Q :=as.factor(pTau_Q)]

vdata %>%
  ggplot( aes(x=pTau_Q, y=IL8.1, fill=Phenotype)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=1, alpha=0.9) +
    theme_ipsum() +
    theme(
      legend.position="right",
      plot.title = element_text(size=11)
    ) +
    ggtitle("IL8-1") +
    xlab("") +
  stat_compare_means(aes(group = Phenotype), label = "p.format")


# do this on discovery/test set 
test_model_matrix <- fread('test_set_all_data.csv')
test_model_matrix[, pTau_Q := ifelse(ptau > 2.6, "T+", "T-")]
test_model_matrix[ , pTau_Q :=as.factor(pTau_Q)]

test_model_matrix %>%
  ggplot( aes(x=pTau_Q, y=IL8.1, fill=Phenotype)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=1, alpha=0.9) +
    theme_ipsum() +
    theme(
      legend.position="right",
      plot.title = element_text(size=11)
    ) +
    ggtitle("IL8-1") +
    xlab("") +
  stat_compare_means(aes(group = Phenotype), label = "p.format")


# MMP7 discovery 
vdata %>%
  ggplot( aes(x=pTau_Q, y=MMP7, fill=Phenotype)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=1, alpha=0.9) +
    theme_ipsum() +
    theme(
      legend.position="right",
      plot.title = element_text(size=11)
    ) +
    ggtitle("MMP7") +
    xlab("") +
  stat_compare_means(aes(group = Phenotype), label = "p.format")

# test set MMP7
test_model_matrix %>%
  ggplot( aes(x=pTau_Q, y=MMP7, fill=Phenotype)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=1, alpha=0.9) +
    theme_ipsum() +
    theme(
      legend.position="right",
      plot.title = element_text(size=11)
    ) +
    ggtitle("MMP7") +
    xlab("") +
  stat_compare_means(aes(group = Phenotype), label = "p.format")


# St2 discovery 
vdata %>%
  ggplot( aes(x=pTau_Q, y=ST2, fill=Phenotype)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=1, alpha=0.9) +
    theme_ipsum() +
    theme(
      legend.position="right",
      plot.title = element_text(size=11)
    ) +
    ggtitle("ST2") +
    xlab("") +
  stat_compare_means(aes(group = Phenotype), label = "p.format")

# test set ST2
test_model_matrix %>%
  ggplot( aes(x=pTau_Q, y=ST2, fill=Phenotype)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=1, alpha=0.9) +
    theme_ipsum() +
    theme(
      legend.position="right",
      plot.title = element_text(size=11)
    ) +
    ggtitle("ST2") +
    xlab("") +
  stat_compare_means(aes(group = Phenotype), label = "p.format")


```
