---
title: "PCA_on_new_genes"
author: "Sudeshna"
date: "12/05/2022"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load required packages
require(data.table)
require(caret)
require(MLeval)
require(glmnet)
require(ROCR)
require(tidyverse)
require(MethylIT)
require(ComplexHeatmap)
require(circlize)
require(psych)

# directories
#data_dir = file.path("..", "..", "data")
rdir = file.path("..", "..", "results", "PCA")
p_dir = file.path(rdir, "Plots")

# load data
ddPCR_r1<-fread(file.path(rdir,"exRNA_ddPCR_gene_exp_for_PCA_051222.csv"))

```



# Preprocessing
```{r}
# create prediction label
ddPCR_r1[, group := map(strsplit(ID, "-"), ~.x[1]) %>% unlist()]
ddPCR_r1[, group := factor(group, levels = c("UA", "DM1"))]
ddPCR_r1[, .N, by=.(group)]
sel_cols = c("ID", "PCK1", "HPD", "DPYS", "GSTA1", "KRT20", "HPGD", "PIGT", "BIN1", "group")
exp = ddPCR_r1[1:49, sel_cols, with = F]
```

Jan 26, 2023
Added new test set: Data_for_predictive_model/exRNA_ddPCR_gene_exp_pred_model_011823.xlsx

```{r}

# splite data into training and testing by 75:25
set.seed(123)
train_idex = createDataPartition(exp$group, p = .75, list = FALSE)
train = exp[train_idex, ]
test = exp[-train_idex, ]
fwrite(test, "test-12-donors.csv")

test<-fread(file.path("..","..","results","Data_for_predictive_model","test-34-donors.csv"))
test[, group := factor(group, levels = c("UA", "DM1"))]
```

```{r}
# normalization
# non-numeric predictors are allowed but will be ignored.
# method = "zv" identifies numeric predictor columns with a single value 
# (i.e. having zero variance) and excludes them from further calculations. 
train_proc = preProcess(train, method = c("center", "scale", "zv"), verbose = TRUE)
train_n = predict(train_proc, train)
head(train_n)

# Define training control
set.seed(123)
train.control = trainControl(method = "repeatedcv", 
                             number = 5, repeats = 10,
                             savePredictions = T,
                             classProbs = T)

```

Model with 8 genes
```{r}
pca.logist.model = train(group ~ PCK1 + HPD + DPYS + GSTA1 + PIGT + BIN1 + HPGD + KRT20, 
                         data = train_n, 
                         method = "glm",
                         trControl = train.control,
                         family = binomial())

print(pca.logist.model)

require(pROC)
p = plot.roc(pca.logist.model$pred$obs,
         pca.logist.model$pred$DM1)

plot(varImp(pca.logist.model), main="Variable Importance")
ruc_dt = data.table(
  false_positive_rate = 1 - p$specificities,
  true_positive_rate = p$sensitivities
 
)
ggplot(ruc_dt, aes(x = false_positive_rate, y = true_positive_rate)) +
  geom_point()
ruc_dt<-ruc_dt[order(false_positive_rate, true_positive_rate),]

fwrite(ruc_dt, file = file.path("..", "..", "results", "PCA", "train_ROC_8_genes_wo_PC.csv"))


# test data
test_pca = predict(
  preProcess(test, method = c("center", "scale", "zv"), verbose = TRUE),
  test
)     
test_predict = predict(pca.logist.model, test_pca)

## confusionMatrix
confusionMatrix(reference = test$group, data = test_predict, 
                mode='everything', positive='DM1')


```


# Exploration
```{r}
# feature correlation analysis
train_cor = cor(train_n[, c(-1, -10)])
train_cor %>% 
  Heatmap(
    col = colorRamp2(c(-1, 0, 1), c("#0C6291", "#FBFEF9", "#A63446")),
    cell_fun = function(j, i, x, y, width, height, fill) {
        grid.text(sprintf("%.1f", train_cor[i, j]), x, y, gp = gpar(fontsize = 10))
})

pairs.panels(train_n[, c(-1, -6)],
             gap = 0,
             bg = c("red", "blue")[train_n$group],
             pch=21)


# 
# require(AppliedPredictiveModeling)
# require(lattice)
# transparentTheme(trans = .4)
# caret::featurePlot(x = ddPCR_n[, 2:6], 
#             y = ddPCR_n$group,
#             plot = "pairs")
```



# Logistic regression using Principal Components from PCA as predictor variables
1. Principal components (PCs) are estimated from the predictor variables provided as input data. 
2. Next, the individual coordinates in the selected PCs are used as predictors in the logistic regresson.

reference: https://github.com/genomaths/MethylIT
```{r}
# step by step
# PCA
set.seed(1234)
train_pca1 = predict(
  preProcess(train, method = c("center", "scale", "pca"), verbose = TRUE),
  train
)   
View(train_pca1)

```

```{r}

require(factoextra)
train_pca2 = prcomp(train[, c(-1, -10)], scale = TRUE, center = TRUE)
train_var = get_pca_var(train_pca2)
#view(train_var)
train_var$contrib
#view(train_var$contrib)

```

```{r}

train_var_dt = train_var$contrib[, 1:4] %>% as.data.frame()
head(train_var_dt)
setDT(train_var_dt)

```

```{r}
train_var_dt[, gene_exp := rownames(train_var$contrib)]
train_var_dt = melt(train_var_dt, id.vars = "gene_exp", 
                    variable.name = "PC", value.name = "contribution")
train_var_dt
```

```{r}
require(ggthemes)
ggplot(train_var_dt, aes(x = PC, y = gene_exp)) +
  geom_point(aes(fill = contribution, size = contribution),
             colour="black",pch=21) +
  theme_few() +
  scale_fill_gradient(low = "yellow", high = "red", na.value = NA) +
  scale_size(range = c(1,10))

```

```{r}

require(cluster)
require(ggfortify)
ggplot(data = train_pca1, aes(x = PC1, y = PC2, col = group, group = group)) + 
       geom_point() + 
       stat_ellipse(type = "norm") +
       theme(legend.position='none') +
       theme_minimal() +
  labs(title = "PC1 vs PC2")
```

```{r}
fwrite(train_pca1, file.path("..", "..", "results", "PCA", "train_pca_8_genes.csv"))
```

```{r}
# ggplot(data = train_pca, aes(x = PC3, y = PC4, col = group, group = group)) + 
#        geom_point() + 
#        stat_ellipse(type = "norm") +
#        theme(legend.position='none') +
#        theme_minimal() +
#   labs(title = "PC3 vs PC4")
```

```{r}
# Define training control
set.seed(123)
train.control = trainControl(method = "repeatedcv", 
                             number = 5, repeats = 3,
                             savePredictions = T,
                             classProbs = T)

```

```{r}

# Train the model
# pca.logist.model = train(group ~ PC1 + PC2, 
#                          data = train_pca1, 
#                          method = "glm",
#                          trControl = train.control,
#                          family = binomial())

```

```{r}
# require(pROC)
# p = plot.roc(pca.logist.model$pred$obs,
#          pca.logist.model$pred$DM1)
# ruc_dt = data.table(
#   true_positive_rate = p$sensitivities,
#   false_positive_rate = 1 - p$specificities
# )
# ggplot(ruc_dt, aes(x = false_positive_rate, y = true_positive_rate)) +
#   geom_point()
# ruc_dt
```

 Final model was with 8 genes and PC1 only
 
```{r}

pca.logist.model_pc1 = train(group ~ PC1, 
                         data = train_pca1, 
                         method = "glm",
                         trControl = train.control,
                         family = binomial())

p = plot.roc(pca.logist.model_pc1$pred$obs,
         pca.logist.model_pc1$pred$DM1)
ruc_dt = data.table(
  true_positive_rate = p$sensitivities,
  false_positive_rate = 1 - p$specificities
)
ggplot(ruc_dt, aes(x = false_positive_rate, y = true_positive_rate)) +
  geom_point()
ruc_dt
fwrite(ruc_dt, file = file.path("..", "..", "results", "PCA", "train_ROC_8_genes_2023-01-26.csv"))
```


```{r}
# pca.logist.model_pc2 = train(group ~ PC2, 
#                          data = train_pca1, 
#                          method = "glm",
#                          trControl = train.control,
#                          family = binomial())

```

```{r}
# Summarize the results
#print(pca.logist.model)
print(pca.logist.model_pc1) #chosen model
#print(pca.logist.model_pc2)

#plot(pca.logist.model, main="Model Accuracies with MARS")
#plot(pca.logist.model_pc2, main="Model Accuracies with MARS")

plot(varImp(pca.logist.model), main="Variable Importance")

```



```{r}
# test data
test_pca = predict(
  preProcess(test, method = c("center", "scale", "pca"), verbose = TRUE, pcaComp = 4),
  test
)     
#test_predict = predict(pca.logist.model, test_pca)
test_predict_pc1 = predict(pca.logist.model_pc1, test_pca)
#test_predict_pc2 = predict(pca.logist.model_pc2, test_pca)
fwrite(data.frame(id=test$ID, group=test$group, pred=test_predict_pc1), file=file.path( "..", "..", "results","PCA","Predictions_TestSet_34donors_2023-01-26.csv"))
```

```{r}
## confusionMatrix
#confusionMatrix(reference = test$group, data = test_predict, 
#                mode='everything', positive='DM1')
confusionMatrix(reference = test$group, data = test_predict_pc1, 
                mode='everything', positive='DM1')
#confusionMatrix(reference = test$group, data = test_predict_pc2, 
#                mode='everything', positive='DM1')
```


```{r}
train_p = copy(train_n)
colnames(train_p) = gsub(" ", "", colnames(train_p))
gene1 = "PCK1"
gene2 = "GSTA1"
ggplot(train_p, aes_string(x = gene1, y = gene2, color = "group", goup = "group")) +
  geom_point() +
  theme_minimal()
ggsave(file.path(p_dir, "train_PCK1_vs_GSTA1.pdf"), width = 5, height = 3)

```


```{r}
save(pca.logist.model, file = "logistic_regression_using_PCA_model_on_new_genes.Rdata")
save(pca.logist.model_pc1, file = "logistic_regression_using_PCA_PC1_model_on_new_genes.Rdata")

```


```{r}
require(cluster)
require(ggfortify)
ggplot(data = test_pca, aes(x = PC1, y = PC2, col = group, group = group)) + 
       geom_point() + 
       stat_ellipse(type = "norm") +
       theme(legend.position='none') +
       theme_minimal() +
  labs(title = "PC1 vs PC2")
```



```{r}
train_p = copy(train_n)
colnames(train_p) = gsub(" ", "", colnames(train_p))
gene1 = "PCK1"
gene2 = "GSTA1"
ggplot(train_p, aes_string(x = gene1, y = gene2, color = "group", goup = "group")) +
  geom_point() +
  theme_minimal()
ggsave(file.path(p_dir, "train_PCK1_vs_GSTA1.pdf"), width = 5, height = 3)
```





```{r}
save(pca.logist.model, file = "logistic_regression_using_PCA_model_on_all_groups.Rdata")
save(pca.logist.model_pc1, file = "logistic_regression_using_PCA_PC1_model_on_all_groups.Rdata")
```

